---
title: "Breaking Bad Viewership Analysis RR Lab 8"
author: "Assyl Salah"
date: "2025-05-24"
output:
  html_document:
    theme: flatly
    df_print: paged
    toc: true
    toc_depth: 2
    code_folding: hide
    self_contained: true
---

# *Breaking Bad*: A Brief Overview

**Breaking Bad** is an American *crime drama* created by **Vince Gilligan**. It aired on **AMC** from **January 20, 2008** to **September 29, 2013**. The series follows *Walter White*, a high-school chemistry teacher turned meth manufacturer, and his former student *Jesse Pinkman*.

::: {style="text-align:center;"}
<img src="Breaking_Bad_logo.svg" width="300px"/>
:::

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.width = 8,
  fig.height= 4
)
library(rvest)
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(DT)
library(ggplot2)

```

## 1. Data Import & Cleaning

*I begin by **scraping** the episode tables from Wikipedia and performing basic **cleaning** so that all columns align properly.*

```{r data}

# 1) Scrape data from Wiki
url    <- "https://en.wikipedia.org/wiki/List_of_Breaking_Bad_episodes"
page   <- read_html(url)
tables <- page %>% html_nodes("table.wikiepisodetable") %>%
          html_table(header = 2, fill = TRUE)

tables_ep <- keep(tables, ~ "No.overall" %in% names(.x))

tables_clean <- map(tables_ep, function(tbl) {
  names(tbl) <- make.names(names(tbl))
  tbl[]      <- lapply(tbl, as.character)
  tbl
})

df_raw <- bind_rows(tables_clean, .id = "Season") %>%
  mutate(Season = as.integer(Season))

vcol <- grep("viewers", names(df_raw), ignore.case=TRUE, value=TRUE)[1]

# Strip out citations and parse numeric
df_raw <- df_raw %>%
  mutate(
    Viewers = as.numeric(str_remove_all(.data[[vcol]], "\\[.*?\\]"))
  )


# Filter only to seasons 1–5
df <- df_raw %>%
  transmute(
    Season         = Season,
    EpisodeOverall = as.integer(str_extract(No.overall, "\\d+")),
    Title          = Title,
    Viewers        = Viewers
  ) %>%
  arrange(Season, EpisodeOverall) %>%
  filter(Season <= 5)
```

**Quick look at the cleaned data**:

```{r peek}
knitr::kable(head(df), caption="*Table: First 6 episodes of the cleaned dataset*")
```

**Data summary**:

```{r summary-raw}
summary(df)
```













