---
title: "Breaking Bad Viewership Analysis RR Lab 8"
author: "Assyl Salah"
date: "2025-05-24"
output:
  pdf_document:
    toc: true
    toc_depth: '2'
  html_document:
    theme: flatly
    df_print: paged
    toc: true
    toc_depth: 2
    code_folding: hide
    self_contained: true
---

# *Breaking Bad*: A Brief Overview

**Breaking Bad** is an American *crime drama* created by **Vince Gilligan**. It aired on **AMC** from **January 20, 2008** to **September 29, 2013**. The series follows *Walter White*, a high-school chemistry teacher turned meth manufacturer, and his former student *Jesse Pinkman*.

::: {style="text-align:center;"}
<img src="Breaking_Bad_logo.svg" width="300px"/>
:::

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.width = 8,
  fig.height= 4
)
library(rvest)
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(DT)
library(ggplot2)

```

## 1. Data Import & Cleaning

*I begin by **scraping** the episode tables from Wikipedia and performing basic **cleaning** so that all columns align properly.*

```{r data}

# 1) Scrape data from Wiki
url    <- "https://en.wikipedia.org/wiki/List_of_Breaking_Bad_episodes"
page   <- read_html(url)
tables <- page %>% html_nodes("table.wikiepisodetable") %>%
          html_table(header = 2, fill = TRUE)

tables_ep <- keep(tables, ~ "No.overall" %in% names(.x))

tables_clean <- map(tables_ep, function(tbl) {
  names(tbl) <- make.names(names(tbl))
  tbl[]      <- lapply(tbl, as.character)
  tbl
})

df_raw <- bind_rows(tables_clean, .id = "Season") %>%
  mutate(Season = as.integer(Season))

vcol <- grep("viewers", names(df_raw), ignore.case=TRUE, value=TRUE)[1]

# Strip out citations and parse numeric
df_raw <- df_raw %>%
  mutate(
    Viewers = as.numeric(str_remove_all(.data[[vcol]], "\\[.*?\\]"))
  )


# Filter only to seasons 1–5
df <- df_raw %>%
  transmute(
    Season         = Season,
    EpisodeOverall = as.integer(str_extract(No.overall, "\\d+")),
    Title          = Title,
    Viewers        = Viewers
  ) %>%
  arrange(Season, EpisodeOverall) %>%
  filter(Season <= 5)
```

**Quick look at the cleaned data**:

```{r peek}
knitr::kable(head(df), caption="*Table: First 6 episodes of the cleaned dataset*")
```

**Data summary**:

```{r summary-raw}
summary(df)
```

## 2. Summary Statistics

*Next, I compute season‐level summaries: number of episodes and average viewers.*

```{r summsry, echo=FALSE}
season_summary <- df %>%
  group_by(Season) %>%
  summarise(
    Avg.Viewers = round(mean(Viewers, na.rm=TRUE), 2)
  )

datatable(
  season_summary,
  rownames = FALSE,
  caption  = htmltools::tags$caption(
    style = 'caption-side: top; text-align: left;',
    "**Table 1.** Average Viewers per Season"
  ),
  options = list(pageLength = 5, dom = 't')
)
```

## 3. Episode‐Level Trend

*Plot 1 shows **viewership for each episode** across all five seasons. Notice the clustering and spread.*

```{r p1}
p1 <- ggplot(df, aes(x=EpisodeOverall, y=Viewers, color=factor(Season))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Viewers per Episode (All Seasons)",
    x     = "Episode (Overall)",
    y     = "Viewers (millions)",
    color = "Season"
  ) +
  theme_minimal(base_size=13)

ggplotly(p1, tooltip=c("x","y","color")) %>%
  layout(hovermode="closest") %>%
  config(displayModeBar=FALSE)
```

## 4. Season‐Level Trend

*Plot 2 shows the **average** viewership per season, highlighting an overall upward trajectory.*

```{r p2}
p2 <- ggplot(season_summary, aes(x=Season, y=Avg.Viewers)) +
  geom_line(size=1.2, linetype="dashed", color="#205237") +
  geom_point(size=4, color="#205237") +
  labs(
    title = "Average Viewers per Season",
    x     = "Season",
    y     = "Viewers (M)"
  ) +
  theme_minimal(base_size=14)

ggplotly(p2, tooltip=c("x","y")) %>%
  layout(hovermode="x unified") %>%
  config(displayModeBar=FALSE)
```

## 5. Season‐to‐Season Change

*Plot 3 illustrates the **change** in average viewership between consecutive seasons.*

```{r p3}
season_summary <- season_summary %>%
  mutate(Delta = round(Avg.Viewers - lag(Avg.Viewers), 2))

p3 <- ggplot(filter(season_summary, Season > 1),
             aes(x=Season, y=Delta, text=paste0("Δ=",Delta," M"))) +
  geom_col(fill="#2d4d00", alpha=0.8) +
  labs(
    title = "Change in Average Viewers Between Seasons",
    x     = "Season",
    y     = "Δ Viewers (M)"
  ) +
  theme_minimal(base_size=14)

ggplotly(p3, tooltip="text") %>%
  config(displayModeBar=FALSE)
```



## 6. Insights & Conclusion
*Breaking Bad* shows a **clear growth** pattern:

From the modest **1.23 million** average viewers in Season 1, the audience steadily increased each year, climbing to **1.31 million** (Season 2), **1.52 million** (Season 3), and **1.87 million** (Season 4). The largest leap occurs between Seasons 4 and 5, with an increase of **2.45 million** viewers to reach an impressive **4.32 million** average viewers.

This trend reflects how the series gained momentum, drawing in more viewers, culminating in a finale that became a cultural phenomenon.

> **Final Thought:**\
> *Breaking Bad’s* Breaking Bad's escalating viewership mirrors its narrative arc, starting small, intensifying mid run, and exploding at the climax.











